---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
options(str = strOptions(strict.width = "cut"))
```

# worldbank

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/m-muecke/worldbank/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/m-muecke/worldbank/actions/workflows/R-CMD-check.yaml)
[![CRAN status](https://www.r-pkg.org/badges/version/worldbank)](https://CRAN.R-project.org/package=worldbank)
<!-- badges: end -->

## Overview

worlbank provides a simple interface to the following [World Bank](https://datahelpdesk.worldbank.org/knowledgebase/articles/889386-developer-information-overview) APIs:

- [Indicators API v2](https://datahelpdesk.worldbank.org/knowledgebase/articles/889392-about-the-indicators-api-documentation)
- [Poverty and Inequality Platform (PIP) API](https://pip.worldbank.org/api).
- [Finances One API](https://financesone.worldbank.org)

The main difference to other packages is that it's a modern implementation using the [httr2](https://httr2.r-lib.org) package
and supports all available endpoints and parameters.

## Installation

You can install the released version of **worldbank** from [CRAN](https://CRAN.R-project.org) with:

```{r, eval = FALSE}
install.packages("worldbank")
```

And the development version from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("m-muecke/worldbank")
```

## Usage

worldbank functions are prefixed with `wb_` and follow the naming convention of the [World Bank API v2](https://datahelpdesk.worldbank.org/knowledgebase/articles/889392-about-the-indicators-api-documentation).

```{r demo}
library(worldbank)

# filter by specific country
wb_country(c("US", "DE"))

# or fetch all (default)
country <- wb_country()
str(country)


# search for specific indicator
ind <- wb_indicator()
ind <- subset(
  ind,
  grepl("GDP", id, fixed = TRUE) & source_value == "World Development Indicators"
)
str(ind)

# fetch indicator data for specific or all countries (default)
gdp <- wb_country_indicator("NY.GDP.MKTP.CD", c("US", "DE", "FR", "CH", "JP"))
str(gdp)
```

```{r plotting, message = FALSE, echo = FALSE, dpi = 300}
library(ggplot2)

subset(gdp, date >= 1980) |>
  ggplot(aes(x = date, y = value, color = country_name)) +
  geom_line() +
  theme_minimal() +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    panel.grid.major.y = element_line(color = "black", linewidth = 0.2),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text = element_text(color = "black"),
    axis.title = element_blank()
  ) +
  scale_y_continuous(
    labels = scales::label_currency(scale_cut = scales::cut_short_scale())
  ) +
  labs(title = "GDP in Current U.S. Dollars", color = "Country")
```

## Related work

* [WDI](https://github.com/vincentarelbundock/WDI): R package to download World Bank data
* [pipr](https://github.com/worldbank/pipr): R client for the PIP Worldbank API
* [wbstats](https://github.com/gshs-ornl/wbstats): R package for searching and downloading data from the World Bank API
* [wbwdi](https://github.com/tidy-intelligence/r-wbwdi): R package to download World Bank indicator data
