---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# worldbank

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/m-muecke/worldbank/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/m-muecke/worldbank/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of worldbank is to provide a simple interface to the [World Bank API v2](World Bank API v2).
The main difference to other packages is that it's a modern implementation using the [httr2](https://httr2.r-lib.org) package without any other
dependencies.

## Installation
You can install the development version of worldbank from [GitHub](https://github.com/) with:

```{r, eval = FALSE}
# install.packages("pak")
pak::pak("m-muecke/worldbank")
```

## Usage

worldbank functions are prefixed with `wb_` and follow the naming convention of the [World Bank API v2](https://datahelpdesk.worldbank.org/knowledgebase/articles/889392-about-the-indicators-api-documentation).

```{r}
library(worldbank)

# filter by specific country
wb_country(c("US", "DE"))

# or fetch all (default)
wb_country()

# fetch indicator data for specific country
wb_country_indicator("NY.GDP.MKTP.CD", "US")
```

## Other World Bank packages
* [wbstats](https://github.com/gshs-ornl/wbstats)
* [WDI](https://github.com/vincentarelbundock/WDI)
